Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model

### 2021-07-14 SIR_manuscript.tex
small edits
small bib/hyphenation fixes

\input{codes/modeldefs.tex}

@@

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
\newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
\newcommand{\testtarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
\newcommand*\subtxt[1]{_{\textnormal{#1}}}
%% Roman subscript
\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}

@@

\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
\newcommand{\needref}{\textcolor{red}{(REFS?)}}

@@

The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting). In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \ali{\citep{peto2020covid,taipale2020population}}. Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates. We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.

@@

Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
primarily diagnostic (determining infection status for clinical purposes)\ali{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \ali{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).

@@

\jd{Discuss the possibility that people who can't get tests might also isolate. This undercuts our assumptions about $\theta\_w$.}

@@

Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).  This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results. Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment). Slowing test returns increases $\Rnum$ only if test positivity (the proportion of infectives in the tested population), and the magnitude of increase in isolation effectiveness from the waiting to the confirmed-positive compartment, are large enough to outweigh the decrease in isolation of uninfected waiting individuals.\david{That sentence is too long to process easily; I've re-organized slightly to make flow a bit better, but it is awkward to really fix it.}

While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure. First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress. This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections. Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns to be an increasing function of test return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate. Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.

The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g. \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]). This phenomenon occurs mainly because of a peculiar feature of the DFE in our model: we implicitly assume that testing is ongoing before the beginning of the epidemic, so that the starting point of the epidemic invasion is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments: in particular, $S\_u^* = \rho/(\rho + \omega)$. (In reality, testing intensity itself is a dynamic variable that will increase as the epidemic proceeds.) If isolation in the waiting compartment is low ($\theta\_w \ll 1$), then waiting individuals can easily be infected once the epidemic starts. Once infected they can still infect others while in the waiting compartment, and cannot be tested again before they have returned to the untested compartment. Thus, under certain extreme conditions (low $\theta\_w$, high $\theta\_c$, low $\omega$, high $\rho$), $\Rnum$ becomes an increasing function of testing intensity. While we can show that this phenomenon \emph{only} occurs under targeted testing ($w_I > w_S$ ), we have not been able to come up with a verbal explanation of the mechanism (this would require interpreting the complicated expression for $a_3$ in Eq.~\ref{eq:abcd2}). In practice, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$. It depends both on the assumption that regular testing is occurring before the epidemic starts, and on levels of testing that are unrealistically high (at least in large, general-population settings), i.e., at least 10\% of the population being tested per day.

Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$. While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public health practices and testing weights is difficult in other contexts. The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc.. This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic. Future steps should explore mathematically tractable ways to model some of these factors more precisely. For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue). Alternatively, one could make the testing weights depend on the testing intensity or test return rate as suggested above. Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.

Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \ali{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
\ali{the following sentences added to address Ben's comment:} There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics. In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results which suggests, as an outbreak control strategy, that proactively screening asymptomatic individuals and isolating confirmed-positive cases would reduce the spread of the disease. In particular, They explored the effects of testing frequency and Trade-offs between frequency, accuracy and the reporting delay on the epidemic dynamic (more specifically, on the effective reproductive number).
\cite{rogers2021high} simulated a SEIR model with testing and isolation which suggest a combination of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals
a strategy for controlling the disease dynamics in the \covid pandemic. Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracing with the focused on posterior predictions and projections under different testing and tracing scenarios. They conclude that the emergence of a second wave depends almost exclusively on the rate at which immunity is lost and that it is necessary to track the 20\% of asymptomatic individuals in order to control the outbreak. Our modeling approach is novel compared to these works as it mechanistically incorporates the epidemic and testing processes with different levels of testing focus from random to highly targeted. We hope that the present paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.

@@

{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$. Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$. We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}. Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$. This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$. Thus, $\rho$ is reparameterized as 
\begin{equation}
\label{eq:phi}
\rho=\frac{\omega \phi}{\omega+\phi}.
\end{equation}
This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
The derivative is 
\begin{align}
\label{eq:dd3dr}
\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
\end{align}
where
\begin{align}
\label{eq:abcd2}
a_3& = \testtarget \Bigg(
&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
& ~ &&+\Big(
\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
+(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
\Big) \omega^2 \gamma \Bigg), \\ \notag
b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
\Big) ,\\ \notag
c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
(2-\theta\_w) \gamma^2 \theta\_w+
(1+\testtarget) \omega \gamma \theta\_w+
\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
\end{align}
Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$. It is straight forward to show that $a_3\geq 0$ in case of random testing strategy, i.e., $w\_S=w\_I=1$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.

@@

We define $w\_{IS}=\testtarget$. \bmb{shouldn't we use this simplified notation \textbf{throughout} the paper???}
\begin{align}
\label{eq:d_del_wis}
\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
\end{align}
which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.

>>>>

\hyphenation{co-ro-na-vi-rus}
\input{codes/modeldefs.tex}

@@

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
\newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
\newcommand*\subtxt[1]{_{\textnormal{#1}}}
%% Roman subscript
\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}

@@

\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
\newcommand{\needref}{\textcolor{red}{(REFS?)}}

@@

The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting). In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \new{\citep{peto2020covid,taipale2020population}}. Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates. We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.

@@

Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
primarily diagnostic (determining infection status for clinical purposes)\new{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \new{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).

@@

@@

Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).  This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results. Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment). Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.

While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure. First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress. This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections. Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate. Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.

The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g. \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]). This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction. A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE. If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again). If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity. 
We can show that this phenomenon \emph{only} occurs under targeted testing ($w_I > w_S$ ), but we have not yet found a simple explanation of why it cannot occur without it. This phenomenon is also unlikely to occur in the real world. In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).

Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$. While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts. The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc.. This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic. Future steps should explore mathematically tractable ways to model some of these factors more precisely. For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue). Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above. Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.

Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \new{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
\new{There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics. In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results which suggest that proactively screening asymptomatic individuals and isolating confirmed-positive cases would reduce the spread of the disease. 
\cite{rogers2021high} simulated a SEIR model with testing and isolation, and suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
H \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios. They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak. 
Our modeling approach differs from earlier work in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted. We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.}

@@

{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$. Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$. We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}. Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$. This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$. Thus, $\rho$ is reparameterized as 
\begin{equation}
\label{eq:phi}
\rho=\frac{\omega \phi}{\omega+\phi}.
\end{equation}
This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
\begin{align}
\label{eq:dd3dr}
\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
\end{align}
where
\begin{align}
\label{eq:abcd2}
a_3& = \testtarget \Bigg(
&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
& ~ &&+\Big(
\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
+(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
\Big) \omega^2 \gamma \Bigg), \\ \notag
b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
\Big) ,\\ \notag
c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
(2-\theta\_w) \gamma^2 \theta\_w+
(1+\testtarget) \omega \gamma \theta\_w+
\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
\end{align}
Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$. It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.

@@

\begin{align}
\label{eq:d_del_wis}
\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
\end{align}
which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
